---
title: "DATA 621 - Homework 1"
output:
  html_document: default
  word_document: default
date: '2022-09-10'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(faraway)
require(corrplot)
require(MASS)
require(reshape)
require(car)
rm(list = ls())
if(!is.null(dev.list()))dev.off()
```

### Variable Selection

Data exploration revealed the presence of multiple missing values for 6 of the predictors. Table XX shows these predictors and the percentage of missing values relative to the predictor.

```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
|Variable|Description|Missing Data (%)|
|---------------|:-------------:|-------------:|
|`TEAM_BATTING_SO`|Strikeouts by batters|5|
|`TEAM_BASERUN_SB`|Stolen bases|6|
|`TEAM_BASERUN_CS`|Caught stealing|34|
|`TEAM_BATTING_HBP`|Batters hit by pitch (get a free base)|92|
|`TEAM_PITCHING_SO`|Strikeouts by pitchers|4|
|`TEAM_FIELDING_DP`|Double Plays|12|
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

It is important to note that stolen bases and caught stealing were not officially tracked until 1887. Records in the dataset date back to 1871; which implies some of the missing data for `TEAM_BATTING_SB` and `TEAM_BASERUN_CS` could be from 1871 to 1886.

### Outliers

Baseball Almanac shows that the record for team batter hits in a season is 1,783 for the Philadelphia Phillies in 1930. The dataset contains several observations where `TEAM_BATTING_H`, the base hits by batters, is above 1,783. Therefore, observations where `TEAM_BATTING_H` is above 1, 783 were removed. Several other variables with similar outliers are documented in Table XX shown below. Observations exceeding the records were removed and each record is by a seasonal basis.

```{r table2, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
|Variable|Indication of Outliers|
|---------------|:-------------:|
|`TEAM_BATTING_H`|Record for team batter hits is 1,783|
|`TEAM_PITCHING_H`|Record for team batter hits is 1,783|
|`TEAM_PITCHING_SO`|Record for strikeouts is 1,595|
|`TEAM_FIELDING_E`|Record for errors is 886|
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```


### Works Cited
Stolen bases: https://en.wikipedia.org/wiki/Stolen_base

Hits records for MLB teams: https://www.baseball-almanac.com/recbooks/hits_records_mlb_teams.shtml

Strikeout records set by MLB teams:
https://www.baseball-almanac.com/recbooks/rb_strike2.shtml

Team errors in a season:
https://www.baseball-fever.com/forum/general-baseball/statistics-analysis-sabermetrics/2403-team-errors-in-a-season