---
title: "Data 621 HW 1"
author: "Krutika Patel"
date: '2022-09-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(olsrr)
library(GGally)
```


# moneyball-evaluation-data.csv
# moneyball-training-data.csv

# 1 Data Exploration

```{r Load Data}
#Local

eval <- as.data.frame(read.csv('moneyball-evaluation-data.csv'))
train <- as.data.frame(read.csv('moneyball-training-data.csv'))
```

```{r Data Summary}
# Column Names
colnames(train)

# Data Summary
summary(train)

# First 5 rows of eval
head(train)

# NA's by Column
sapply(train, function(x) sum(is.na(x)))

# Select Necessary Data
# Remove index col
data <- train[-1]
```

```{r Mean and Standard Deviation}
# Mean and SD
m_sd <- data.frame(matrix(data=NA, 1,2))
colnames(m_sd) <- c('Mean', 'Standard_Deviation')
for(col in data)
{
  m_sd[nrow(m_sd) + 1,] = c(mean(col, na.rm = TRUE), sd(col, na.rm = TRUE))
}
m_sd <- m_sd[-1,]
m_sd
```

```{r Graphics}
pairs(data)
```

```{r}
plot <- melt(train, "TARGET_WINS")
ggplot(data = plot, aes(value, TARGET_WINS)) +
  geom_point() +
  facet_wrap(.~variable, scales = "free") +
  geom_smooth(method = "lm")
```


# 2 Data Preperation

```{r}
# Simplest Data preperation
# Replacing NA's with mean/median

for(col in train)
{
  col[is.na(col)] <- mean(col) #median(col)
}

```

NA Approach

Columns with NAs

TEAM_BATTING_SO 102: Strikeouts by pitchers 

TEAM_BASERUN_SB 131: Stolen bases 

TEAM_BASERUN_CS 772: Caught stealing 

TEAM_BATTING_HBP 2085: Batters hit by pitch (get a free base)

TEAM_PITCHING_SO 102: Strikeouts by batters 

TEAM_FIELDING_DP 286: Double Plays 


# 3 Build Models`

```{r}
# Simple Linear Regression Model

lm <- lm(TARGET_WINS ~ TEAM_BATTING_H, data = train)
p_tw <- data.frame('Predicted Target Wins' = predict(lm,eval))
p_tw
colnames(train)
```

```{r All possible models}
# Run all possible models using library olsrr
model <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS +  TEAM_BATTING_HBP + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP, data = train)

all_models <- olsrr::ols_step_all_possible(model)

```

```{r}
all_models
```
Double check lm model to see if we get the same adj r^2

# 4 Select Models

```{r Select top n models}

top_n_models <- all_models %>%
  arrange(desc(adjr)) %>%
  slice(1:3)

top_n_models$predictors[1]
```

