---
title: "Final_3"
author: "Krutika Patel"
date: '2022-09-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval-load, eval=FALSE}
#import dataset: moneyball_evaluation_data
eval <- read_csv("moneyball-evaluation-data.csv")
```

```{r train-load, eval=FALSE}
#import dataset: moneyball_training_data
train <- read_csv("moneyball-training-data.csv")
```

```{r}

# run this for training data

new_cols <- c("index", "target", "bat_h", "bat_2b", "bat_3b", "bat_hr", "bat_bb", "bat_so", "bas_sb", "bas_cs", "bat_hbp", "p_h", "p_hr", "p_bb", "p_so", "f_e", "f_dp"
)

colnames(train) <- new_cols
```


```{r}

# run this for evaluation data

new_cols <- c("index", "bat_h", "bat_2b", "bat_3b", "bat_hr", "bat_bb", "bat_so", "bas_sb", "bas_cs", "bat_hbp", "p_h", "p_hr", "p_bb", "p_so", "f_e", "f_dp"
)

colnames(eval) <- new_cols
```

```{r}
data <- train
```


### Examine base model, no transformations, no engineering

```{r}
mdl <- lm(target~., data=data)
summary(mdl)
```

### Create Saber Feature

Saber Model

Sabermetrics has become the rage in baseball, actually popularized by Billy Beane and the data set we are exploring. As a result, we built a model that centers around one of these advance analytics known as BsR or base runs. This statistic (designed by David Smyth in the 1990’s) estimates the amount of runs a team SHOULD score, adding an intriguing element to a data set which does not include runs (see http://tangotiger.net/wiki_archive/Base_Runs.html for more information). The formula For BsR is as follows:

BSR = A*B/(B+A)+C where:

A = TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_BB

B = 1.02(1.4TEAM_TOTAL_BASES -0.6TEAM_BATTING_H + 0.1TEAM_BATTING_BB)

C = TEAM_BATTING_HR

Since we eliminate the value of TEAM_BATTING_H, we sum up singles, doubles, triples and home runs in the actual code, and the approach for TEAM_TOTAL_BASES is described in model 2. The data for BSR exhibit a fairly normal distribution.


```{r}
data$bat_1b <- data$bat_h - data$bat_2b - data$bat_3b - data$bat_hr
data$total_bases <- data$bat_1b + 2*data$bat_2b + 3*data$bat_3b + 4*data$bat_hr


A <- data$bat_h
B <- 1.02*(1.4*data$total_bases -0.6*data$bat_h + 0.1*data$bat_bb)
C <- data$bat_hr

data$saber <- A*B/(B+A)+C
```

```{r}
mdl_inc_saber <- lm(target~., data=data)
summary(model)
```

## First Model - highest correlating features (above 0.2)

- total_bases: 0.423
- saber: 0.395
- bat_h: 0.353
- bat_2b: 0.289
- bat_bb: 0.233
- bat_1b: 0.217
- p_bb: 0.204

### Top features without engineering

```{r}
mdl_tf <- lm(target~ bat_h + bat_2b + bat_bb + bat_1b + p_bb, data=data)
summary(model)
```


### Remove correlating features from feature engineering (just saber)

```{r}
model <- lm(target~ saber, data=data)
summary(model)
```

### Top features including engineering

```{r}
mdl_tf_fe <- lm(target~ total_bases + saber + bat_h + bat_2b + bat_bb + bat_1b + p_bb, data=data)
summary(model)
```

### Step AIC

```{r}
mdl_step.model <- stepAIC(mdl, direction = "both", 
                      trace = FALSE)
summary(mdl_step.model)
```

### Square Root Step AIC

```{r}
mdl_sqrt = lm(sqrt(target) ~ .,
               data = data)
summary(mlr2)
```

```{r}
mdl_sqrt_step.model <- stepAIC(mdl, direction = "both", 
                      trace = FALSE)
summary(mlrstep.model)
```

### All models

```{r}
mdl_all <- olsrr::ols_step_all_possible(model)
```


