---
title: "621 Homework 1"
author: "Alec"
date: "9/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r include=FALSE}
library(dplyr)
library(tidyverse)
library(GGally)
library(MASS)
library(reshape2)
```


## Load Data

```{r results=FALSE, message=FALSE}
data <- read_csv("../data/moneyball-training-data.csv")
```

## Part 1 - Data Exploration

```{r}
# There are 2276 observations, with 16 features and 1 target variable

dim(data)
```

There are 2276 observations, with 16 features and 1 target variable

```{r}
# All fields are of type double.
head(data)
```


```{r}
summary(data)
```


### Check for Missing Values

```{r}
sapply(data, function(x) sum(is.na(x))/dim(data)[1])
```

In total there are 6 columns with missing values:
- Strikeouts by batters (5%)
- Stolen bases (6%)
- Caught stealing (34%)
- Batter hit by pitch (92%)
- Strikeouts by pitchers (4%)
- Double plays (12%)

We will impute these columns using each's respective median value. We will discard "Batter hit by pitch" due to 92% of entries missing.


```{r}
data$TEAM_BATTING_SO[is.na(data$TEAM_BATTING_SO)] <- median(data$TEAM_BATTING_SO, na.rm = T)
data$TEAM_BASERUN_SB[is.na(data$TEAM_BASERUN_SB)] <- median(data$TEAM_BASERUN_SB, na.rm = T)
data$TEAM_BASERUN_CS[is.na(data$TEAM_BASERUN_CS)] <- median(data$TEAM_BASERUN_CS, na.rm = T)
data$TEAM_PITCHING_SO[is.na(data$TEAM_PITCHING_SO)] <- median(data$TEAM_PITCHING_SO, na.rm = T)
data$TEAM_FIELDING_DP[is.na(data$TEAM_FIELDING_DP)] <- median(data$TEAM_FIELDING_DP, na.rm = T)
```


```{r}
# remove unused columns

data <- data %>% dplyr::select(!c("INDEX", "TEAM_BATTING_HBP"))
```



```{r}
# check final results after imputation

sapply(data, function(x) sum(is.na(x))/dim(data)[1])
```

Rename the columns for easier reading

```{r}
new_cols <- c("target", "bat_h", "bat_2b", "bat_3b", "bat_hr", "bat_bb", "bat_so", "bas_sb", "bas_cs", "p_h", "p_hr", "p_bb", "p_so", "f_e", "f_dp"
)

colnames(data) <- new_cols
```


Check the distribution of the target variable

```{r}
data %>%
  ggplot() +
  geom_density((aes(x=target)))
```

# get distribution of all features

```{r}
ggplot(stack(data), aes(x = ind, y = values)) +
  geom_boxplot()
```



The target variable is normally distributed. Let's check the distributions and correlations of all other features.

```{r}
ggpairs(data, upper = list(continuous = wrap("cor", size=2)))
```

Top correlating features:
- bat_h: 0.389
- bat_2b: 0.289
- bat_bb: 0.233
- p_hr: 0.189
- bat_hr: 0.176
- f_e: -0.176
- bat_3b: 0.143

Looking at the above, more than anything, features related to batting are the highest correlating with the target variable

Things we will want to address:
- deal with skewed features:
  - bat_3b
  - bas_sb
  - bas_cs
  - p_h
  - p_bb
  - p_so
  - f_e
- deal with bimodal features:
  - bat_hr
  - p_hr


# Data Preparation

```{r}
summary(data$bat_3b)
```


# remove outliers

```{r}
detect_outlier <- function(x) {
   
    # calculate first quantile
    Quantile1 <- quantile(x, probs=.25)
   
    # calculate third quantile
    Quantile3 <- quantile(x, probs=.75)
   
    # calculate inter quartile range
    IQR = Quantile3-Quantile1
   
    # return true or false
    x > Quantile3 + (IQR*1.5) | x < Quantile1 - (IQR*1.5)
}
 
# create remove outlier function
remove_outlier <- function(dataframe,
                            columns=names(dataframe)) {
   
    # for loop to traverse in columns vector
    for (col in columns) {
       
        # remove observation if it satisfies outlier function
        dataframe <- dataframe[!detect_outlier(dataframe[[col]]), ]
    }
   
    # return dataframe
    print("Remove outliers")
    print(dataframe)
}
```



Apply log transformation for all skewed features

```{r}
data$bat_3b <- log(data$bat_3b + 1)
data$bas_sb <- log(data$bas_sb + 1)
data$bas_cs <- log(data$bas_cs + 1)
data$p_h <- log(data$p_h + 1)
data$p_bb <- log(data$p_bb + 1)
data$p_so <- log(data$p_so + 1)
data$f_e <- log(data$f_e + 1)

```


```{r}
ggpairs(data, upper = list(continuous = wrap("cor", size=2)))
```


# Build Models

## First Model - highest correlating features

- bat_h: 0.389
- bat_2b: 0.289
- bat_bb: 0.233
- p_hr: 0.189
- bat_hr: 0.176
- f_e: -0.176
- bat_3b: 0.143

```{r}
model <- lm(target~ bat_h + bat_2b + bat_bb + p_hr + bat_hr + f_e + bat_3b, data=data)
summary(model)
```


```{r}
model <- lm(target~., data=data)

summary(model)
```






