---
title: "Final_1"
author: "Krutika Patel"
date: '2022-09-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Data Exploration

```{r}
suppressPackageStartupMessages(library(tidyverse))
```


# Importing the datasets
```{r eval-load, eval=FALSE}
#import dataset: moneyball_evaluation_data
eval <- read_csv("moneyball-evaluation-data.csv")
```

```{r train-load, eval=FALSE}
#import dataset: moneyball_training_data
train <- read_csv("moneyball-training-data.csv")
```

# Viewing Data
```{r}
head(eval)
```

```{r}
head(train)
```

# Column Names
```{r}
as.data.frame(colnames(eval))
```

```{r}
as.data.frame(colnames(train))
```

# Data Subset
```{r}
#remove/separate index column from dataset
head(eval)
new_eval <- subset(eval, select = -INDEX)
eval_index <- eval$INDEX
```

```{r}
new_train <- subset(train, select = -INDEX)
train_index <- train$INDEX
```

# Summary Statistics
```{r}
summary(train)
```

# NA exploration
```{r}
sapply(new_train, function(x) sum(is.na(x)))
```

```{r}
suppressPackageStartupMessages(library("visdat"))
```

```{r}
new_train  %>%
  summarise_all(list(~is.na(.)))%>%
  pivot_longer(everything(),
               names_to = "variables", values_to="missing") %>%
  count(variables, missing) %>%
  ggplot(aes(y=variables,x=n,fill=missing))+
  geom_col()+
  scale_fill_manual(values=c("skyblue3","gold"))+
  theme(axis.title.y=element_blank()) + theme_classic()

```

```{r}
suppressPackageStartupMessages(library("VIM"))
```

```{r}
aggr_plot <- aggr(new_train, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(new_train), cex.axis=.48, gap=2, ylab=c("Histogram of missing data","Pattern"))
```

```{r}
suppressPackageStartupMessages(library("scales"))
```


```{r}
library("DataExplorer")
```


```{r}
plot_missing(new_train, missing_only = TRUE, geom_label_args = list("size" = 3, "label.padding" = unit(0.1, "lines")))
```

# Correlation Plots
```{r}
suppressPackageStartupMessages(library(corrplot))
```


```{r}
#correlation matrix
#using only pairwise-complete observations to avoid NA values
M <- cor(new_train, use="pairwise.complete.obs")

#visualizing correlogram
corrplot(M, method="circle", tl.col = "black", tl.cex = 0.6, tl.srt = 70)

#as pie
corrplot(M, method = "pie", tl.col = "black", tl.cex = 0.6, tl.srt = 70 )

#as color
corrplot(M, method = "color", tl.col = "black", tl.cex = 0.6, tl.srt = 70)

#as number
corrplot(M, method = "number", tl.col = "black", tl.cex = 0.6, tl.srt = 70)
```

# Boxplots

Apply the `pivot_longer function` to reshape some of the columns of our data from wide to long format:
```{r}
#Reshape data frame
train_long <- pivot_longer(new_train,
                          c("TARGET_WINS", "TEAM_BATTING_H", "TEAM_BATTING_2B", "TEAM_BATTING_3B", "TEAM_BATTING_HR", "TEAM_BATTING_BB", "TEAM_BATTING_SO", "TEAM_BASERUN_SB", "TEAM_BASERUN_CS", "TEAM_BATTING_HBP", "TEAM_PITCHING_H", "TEAM_PITCHING_HR", "TEAM_PITCHING_BB", "TEAM_PITCHING_SO", "TEAM_FIELDING_E", "TEAM_FIELDING_DP"))
```


Apply the ggplot and geom_boxplot functions to the `train_long` data to visualize each of the selected columns in a side-by-side boxplot graphic:
```{r message=FALSE, warning=FALSE}
#Draw boxplots
#rows containing non-finite value (stat_boxplot) removed NA rows
ggplot(train_long, 
       aes(x = log(value),
           fill = name)) +
  geom_boxplot()
```

# Histograms

Creating a histogram for each of our columns, and using the `facet_wrap function` to separate each column in its own plotting panel:

```{r}
#Draw histograms
ggplot(train_long, 
       aes(x = log(value))) +
  geom_histogram(bins = 15) +
  facet_wrap(name ~ ., scales = "free")
```

